---
title: "Codes for using ANovas and making charts"
output: html_notebook
---


# Coding

### Install packages

```{r}
# This package teaches R how to import SPSS files
if (!require(haven)){
  install.packages("haven", dependencies = TRUE)
  require(haven)
}

# This package has a million useful functions for tidying, cleaning, and reshaping data
if (!require(tidyverse)){
  install.packages("tidyverse", dependencies = TRUE)
  require(tidyverse)
}

# This package has functions for doing Type 3 Sum of Squares ANOVAs
if (!require(afex)){
  install.packages("afex", dependencies = TRUE)
  require(afex)
}

# This package calculates effects sizes for t-tests
if (!require(effsize)){
  install.packages("effsize", dependencies = TRUE)
  require(effsize)
}

# This package has helpful functions for reporting statistics in a manuscript
if (!require(papaja)){
  install.packages("papaja", dependencies = TRUE)
  require(papaja)
}
```

### Import the data file
```{r}
dataset <- read_sav("https://osf.io/re9yh/download")
```

### View the data file
```{r}
View(dataset)
```
=
### Remove duplicate rows
```{r}
dataset %>% #start with dataset
  distinct(Subject, .keep_all = TRUE) -> dataset.unique
```


```{r}
(dataset.clean.aov <- aov_ez(id = "SubjectID", dv = "CandidateRace", data = dataset.clean, between=c("Q17))
```


```{r}
(dataset.unique %>% # start with the dataset 
  group_by(EC) %>% # group by the training goal variable
  summarise(mean = mean(RaceDFeb), # get the mean of the DV in each group
            SD = sd(RaceDFeb)) -> trainingGoalDescriptives) # get the SD of the DV in each group
```

```{r}
aov_recall <- apa_print(dataset.unique.aov, es = "pes", in_paren=FALSE) # es = pes is how we get the partial eta squared reproted
```
And we will use this helper function to round variables to 2 decimal places
```{r}
twoDecimalFormat <- function(val){
  format(round(val, 2), nsmall = 2)
}
```

USE AOV RECALL TO WRITE RESULTS

### Figure
#### Create a single variable with four levels that correspond to the four conditions of the experiment.
```{r}
dataset.unique %>% #start with dataset
  unite("TrainingTypebyGoal", c("TrainTyp", "EC")) %>% # combine the 2 IV columns
  mutate(TrainingTypebyGoal = if_else(str_detect(TrainingTypebyGoal, "1_0"), "Iden Training, Reduce Bias", TrainingTypebyGoal), 
         TrainingTypebyGoal = if_else(str_detect(TrainingTypebyGoal, "1_1"), "Iden Training, Maintain Bias", TrainingTypebyGoal),
         TrainingTypebyGoal = if_else(str_detect(TrainingTypebyGoal, "2_0"), "Attitude Training, Reduce Bias", TrainingTypebyGoal),
         TrainingTypebyGoal = if_else(str_detect(TrainingTypebyGoal, "2_1"), "Attitude Training, Maintain Bias", TrainingTypebyGoal)) -> dataset.unique.plotDF
```
#### Create bar graph
```{r}
ggplot(dataset.unique.plotDF, aes(x = factor(TrainingTypebyGoal), y=RaceDFeb)) + 
  stat_summary(fun.y="mean", geom="bar", width=.4, color="black", fill="white") + 
  theme_minimal() + 
  theme(panel.grid = element_blank(), axis.line.y = element_line(color="black", size = .5), axis.text.x  = element_text(angle=90, vjust=0.5)) + 
  geom_hline(yintercept=0, color="black", size=.5) + 
  labs(x = "", y = "Implicit Prejudice") + 
  stat_summary(fun.data = "mean_cl_boot", geom="errorbar", width=.05) 
```

## 2x2 ANOVA: one IV between, one IV within
If you're ANOVA has at least one within subjects variable then you need to reshape your dataset from wide to long
```{r}
dataset.unique %>% # start with dataset
  gather(ImplicitBias, value, -Subject, -TrainTyp, -EC, -IATOrder) -> dataset.unique.long # reshape from wide to long
```

### ANOVA code

```{r}
(dataset.unique.rm <- aov_ez("Subject","value", dataset.unique.long, between=c("EC"), within=c("ImplicitBias")))
```

# If the interaction was significant in either ANOVA, this is what we would do

### Estimate marginal means
```{r}
(ls.within <- lsmeans(dataset.unique.rm, c("Q17"), by = c("CandidateRace")))
```
### Test simple Effects
```{r}
(simple.effects.rm <- update(pairs(ls.within), by=NULL, adjust = "none"))
```
### Calculate the effect size associated with each t-test.


#### First, we'll calculate the cohen's d for when implicit prejudice is the DV
```{r}
(dataset.unique.long %>%
  filter(ImplicitBias == "RaceDFeb") %>%
   with(effsize::cohen.d(value ~ EC, paired = FALSE, hedges.correction=FALSE, na.rm=TRUE)) -> RaceDFebEffSize)
```
#### Now, we'll calculate the cohen's d for when identification is the DV
```{r}
(dataset.unique.long %>%
  filter(ImplicitBias == "IdenDFeb") %>%
   with(effsize::cohen.d(value ~ EC, paired = FALSE, hedges.correction=FALSE, na.rm=TRUE)) -> IdenDFebEffSize)
```
