---
title: "Project Group 3 Analysis Notebook"
output:
  html_notebook:
  theme: paper
  toc: yes
  toc_float: yes
---


```{r}
dataset <- read_sav("Hiring_Decisions.sav")
```




# Clean Dataset
## Remove practice participants and participants that didn't finish

```{r}
dataset %>%
    filter(Status == 0) %>%
    filter(Progress > 90)%>%
    filter(Duration__in_seconds_> 120) -> dataset.clean
```

# Create Independent Variables
## Making a candidate Race IV
1= Black, 2= White
```{r}
dataset.clean %>%
mutate(CandidateRace =  coalesce(FL_14_DO_C3_Black_LowQualification,
        FL_14_DO_C2_Black_MediumQualification,
        FL_14_DO_C1_Black_HighQualification))%>%
mutate(CandidateRace = replace_na(CandidateRace, 2)) -> dataset.clean
  
```

## Make Candidate Qualifications IV
1 = low, 2 = ambiguous, 3 = high
```{r}
dataset.clean %>%
  mutate(CandidateQualifications = coalesce(
    FL_14_DO_C1_Black_HighQualification, 
    FL_14_DO_C4_White_HighQualification)) %>%
  mutate(CandidateQualifications = case_when(
    CandidateQualifications == 1 ~ 3,
    TRUE~CandidateQualifications)) %>% 
  mutate(CandidateQualifications = coalesce(
    CandidateQualifications, 
    FL_14_DO_C3_Black_LowQualification, 
    FL_14_DO_C6_White_LowQualification)) %>%
  mutate(CandidateQualifications = replace_na(
    CandidateQualifications, 2)) -> dataset.clean
```
# Dependent Variables
```{r}
dataset.clean %>%
  select(Q17, Q18) -> DataFrame

print(dfSummary(DataFrame, graph.magnif = .75), method = 'render')
```

## Would you recommend hiring this applicant
```{r}
(aov_ez (id = "ResponseId",
        dv = "Q17",
        data = dataset.clean,
        between = c("CandidateRace", "CandidateQualifications")) -> dataset.clean17.aov)

```

### Summarize main effect of candidate qualifications
```{r}
dataset.clean %>%
  group_by(CandidateQualifications) %>%
  summarise(mean(Q17, na.rm=TRUE),
  sd = sd(Q17, na.rm= TRUE))
```
```{r}
dataset.clean.aov %>%
  lsmeans(c("CandidateQualifications")) %>%
  pairs()
```

### Summarize Main Effect of Candidate Race
```{r}
dataset.clean %>%
  group_by(CandidateRace) %>%
  summarise(mean(Q17, na.rm=TRUE),
  sd = sd(Q17, na.rm= TRUE))
```
### Summarize all Conditions
```{r}
dataset.clean %>%
  group_by(CandidateRace, CandidateQualifications) %>%
  summarise(mean(Q17, na.rm=TRUE),
  sd = sd(Q17, na.rm= TRUE))
```

```{r}
dataset.clean %>%
  with(t.test(Q18, CandidateQualifications))
```

```{r}
dataset.clean %>% #start with dataset
  unite("CandidateQualificationsbyRace", c("Q17")) %>% # combine the 2 IV columns
  mutate(CandidateQualificationsbyRace = if_else(str_detect(CandidateQualificationsbyRace, "0"), "Low, Black", Q18), 
         CandidateQualificationsbyRace = if_else(str_detect(CandidateQualificationsbyRace, "2"), "Low, White", Q18),
         CandidateQualificationsbyRace = if_else(str_detect(CandidateQualificationsbyRace, "3"), "Medium, Black", Q18),
         CandidateQualificationsbyRace = if_else(str_detect(CandidateQualificationsbyRace, "4"), "Medium, White", Q18), 
         CandidateQualificationsbyRace = if_else(str_detect(CandidateQualificationsbyRace, "5"), "High, Black", Q18),
         CandidateQualificationsbyRace = if_else(str_detect(CandidateQualificationsbyRace, "6"), "High, White", CandidateQualificationsbyRace)) -> dataset.clean.plotDF
```

```{r}
ggplot(dataset.clean.plotDF, aes(x = factor(CandidateQualificationsbyRace), y= Q18)) + 
  stat_summary(fun.y="mean", geom="bar", width=.4, color="black", fill="white") + 
  theme_minimal() + 
  theme(panel.grid = element_blank(), axis.line.y = element_line(color="black", size = .5), axis.text.x  = element_text(angle=90, vjust=0.5)) + 
  geom_hline(yintercept=0, color="black", size=.5) + 
  labs(x = "", y = "Preference") + 
  stat_summary(fun.data = "mean_cl_boot", geom="errorbar", width=.05)
```

```{r}
aov_recall <- apa_print(dataset.clean.aov, es = "pes", in_paren=FALSE)
```

```{r}
twoDecimalFormat <- function(val){
  format(round(val, 2), nsmall = 2)
}
```

```{r}
(dataset.clean %>% # start with the dataset 
  group_by(CandidateQualifications) %>% # group by the training goal variable
  summarise(mean = mean(Q17), # get the mean of the DV in each group
            SD = sd(Q17)) -> CandidateQualificationsDescriptives)
```

```{r}
effect.size(CandidateQualifications, Q17, pooled = TRUE, conf.level = 0.95)

```


To examine the effect of candidate race and candidate qualifications on hiring decisions and hiring recommendations, we conducted a 2 $x$ 2 subjects ANOVA with hiring decisions as the dependent variable. Contrary to what we expected, the interaction between candidate race and hiring decisions, `r aov_recall$full_result$CandidateRace`, was not significant. Along with this, surprisingly, the interaction between candidate race and qualifications with hiring decisions was also not significant, `r aov_recall$full_result$CandidateRace_CandidateQualifications`. Interestingly, the interaction between candidate qualifications and hiring decisions was significant, `r aov_recall$full_result$CandidateQualifications`, indicating that participants were impartial to the race of their applicant, and put more emphasis on applicant qualifications, (*M* = `r twoDecimalFormat(CandidateQualificationsDescriptives$mean[1])`, *SD* = `r twoDecimalFormat(CandidateQualificationsDescriptives$SD[1])`).